### 2017-09-26 article.tex
basic format
Empty format

\documentclass{article}

\usepackage{amsmath}

%opening
\title{ISF Grant}
\author{}

\begin{document}

\maketitle

\begin{abstract}
	a
\end{abstract}

\bibliographystyle{abbrv}
\bibliography{references}

\end{document}

### 2017-09-26 article.tex
introduction

@@
 
+\section{Introduction}
+
+Traditional computer architectures containing both a volatile main memory alongside a non-volatile secondary storage, have led to a separation of program state into operational data stored using in-memory data structures, and recovery data stored using sequential on-disk structures such as transaction logs.
+A system-wide failure, such as a power outage, cause a full lost of the in-memory data structures, which must be reconstructed entirely from the recovery data, making the software system temporarily unavailable.
+As a result, the design of in-memory data structures emphasizes disposable constructs optimized for parallelism, in contrast to the structures that hold recovery data, which cannot be discarded upon a failure and which beneft less from parallelism since their performance is limited by the secondary storage.
+
+Current industry projections indicate that non-volatile main memory (NVRAM) will become commonplace over the next few years.
+While the availability of NVRAM suggests the possibility of keeping persistent data in main memory (not just in the file system), the fact that recent updates to registers and cache may be lost during a power failure means that the data in main memory, if not carefully managed, may not be consistent at recovery time.
+Traditional log-based recovery techniques, although can be applied correctly in such systems, fail to take full advantage of the parallelism enabled by allowing processing cores to access recovery data directly using memory operations rather than slow block transfers from secondary storage.
+As a result, harnessing the performance benefits of NVRAM-based platforms requires a careful rethinking of recovery mechanisms.
+
+The study of shared data structures and concurrent programming dates back over a half century ago.
+1965 Dijkstra's pioneering work presented the concept of concurrent programming [*], followed by a series of seminal papers on inter-process communication, wait-free synchronization, and linearizability [*].
+Recent research has paid close attention to asynchronous models, in which there is no bound on the amount of time a process takes to transition to its next step, or to complete a memory operation.
+This assumption captures the behavior of most modern memory hierarchies, where different media (e.g., L1 cache vs.
+L2 cache vs.
+main memory) incur vastly different and often unpredictable access latencies, as well as the effect of the operating system (e.g., via preemption and interrupts) on the liveness of processes.
+In addition, asynchrony is also related to reliability in a precise way: algorithms that provide non-blocking progress properties (e.g., lock-freedom and wait-freedom) in an asynchronous environment with reliable processes continue to provide the same progress properties if crash failures are introduced.
+Informally speaking, this property holds because a process that crashes permanently at an arbitrary point in the execution of its algorithm is indistinguishable to the other processes from one that is merely very slow.
+
 \bibliographystyle{abbrv}

### 2017-09-26 article.tex
correctness property

@@
 
+\subsection{Correctness and Progress properties}
+Correctness property specifies the requirments for an implementation to be considerd correct.
+Linearizability, presented by Wang and Herliry [*], is the most common correctness property in use for shared memory objects.
+Roughtly speaking, linearizability requires that operations on objects appear to take effect instantaneously in some serial order, while preserving the happens before relation, i.e., if op1 ends before op2 begins, then op1 should precede op2 in the serial order.
+Linearizability has no notion of an aborted or failed operation, and requires that a process finish one operation before it invokes the next.
+The criteria according to which implementations that support recovered processes are correct remain unclear.
+
 \bibliographystyle{abbrv}

### 2017-09-27 article.tex
RR-linearizability

--- 2017-09-26 article.tex
@@
 A system-wide failure, such as a power outage, cause a full lost of the in-memory data structures, which must be reconstructed entirely from the recovery data, making the software system temporarily unavailable.
-As a result, the design of in-memory data structures emphasizes disposable constructs optimized for parallelism, in contrast to the structures that hold recovery data, which cannot be discarded upon a failure and which beneft less from parallelism since their performance is limited by the secondary storage.
+As a result, the design of in-memory data structures emphasise disposable constructs optimized for parallelism, in contrast to the structures that hold recovery data, which cannot be discarded upon a failure and which benefit less from parallelism since their performance is limited by the secondary storage.
 
@@
 Recent research has paid close attention to asynchronous models, in which there is no bound on the amount of time a process takes to transition to its next step, or to complete a memory operation.
-This assumption captures the behavior of most modern memory hierarchies, where different media (e.g., L1 cache vs.
+This assumption captures the behaviour of most modern memory hierarchies, where different media (e.g., L1 cache vs.
 L2 cache vs.
@@
 
+\subsection{Correctness properties}
+Correctness property specifies the requirements for an implementation to be considered correct.
+Herlihy and Wing’s linearizability property [*] is widely adopted for shared objects in conventional shared memory models.
+Roughly speaking, linearizability requires that operations on objects appear to take effect instantaneously in some serial order, while preserving the happens before relation, i.e., if $op_1$ ends before $op_2$ begins, then $op_1$ should precede $op_2$ in the serial order.
+Linearizability has no notion of an aborted or failed operation, and requires that a process finish one operation before it invokes the next.
+The criteria according to which implementations that support crash-recovery failures are correct remains unclear.
+
+Aguilera and Frølund have proposed strict linearizability as a safety condition for persistent concurrent objects, which treats the crash of a process as a response, either successful or unsuccessful, to the interrupted operation [*].
+A successful response means that the operation takes effect at some point between its invocation and the crash failure, and an unsuccessful response means that the operation does not take effect at all.
+Strict linearizability preserves both locality and program order.
+However, they proved that it precludes the implementation of some wait-free objects for certain machine models - there is no wait-free implementations of multi-reader single-writer (MRSW) registers from single-reader single-writer (SRSW) registers under strict linearizability.
+
+Guerraoui and Levy have proposed persistent atomicity [*].
+It is similar to strict linearizability, but allows an operation interrupted by a failure to take effect before the subsequent invocation of the same process, possibly after the failure.
+They also proposed transient atomicity, which relaxes this criterion even further and allows an interrupted operation to take effect before the subsequent write response of the same process.
+Both conditions ensure that the state of an object will be consistent in the wake of a crash, but it does not provide locality: correct histories of separate objects, when merged, will not necessarily yield a correct composite history.
+
+Berryhill et al.
+have proposed an alternative, recoverable linearizability [*], which achieves locality but may sacrifice program order after a crash.
+It is a relaxed version of persistent atomicity, which requires the operation to be linearizes or aborts before any subsequent linearization by the pending thread on that same object.
+
+Izraelevitz et al.
+considered a real-world failure model, in which processes are assumed to fail together, as part of a full-system crash.
+Under this model, persistent atomicity and recoverable linearizability are indistinguishable (and thus local).
+The term durable linearizability was used to refer to this merged safety condition under the restricted failure model.
+
+\subsubsection{Recoverable response}
+All the correctness properties discussed above focused on keeping the object or data structure consist in case of a failures.
+However, none of which address the problem of the returned value by the failed operation.
+Consider for example a write operation.
+As the operation is atomic, it obviously satisfies strict linearizability - either the write took effect before the crash failure, or that it did not.
+However, once the process is recovered it has no knowledge which of the two is the case, since the acknowledgement returned by the write is saved to the local memory of the process (part of the cache), which is lost due to the failure.
+Assuming the process is in a middle of executing a bigger operation which it desired to complete once it is recovered, it must know whether the write took effect (in which case it can continue and execute the rest of the code), or that it did not (in which case it needs to rewrite).
+Skipping the write is not an option, as it might cause an undesired behaviour which in turn might violate the requirements of the object.
+
+For this case, we present a different approach called Recoverable response linearizability (RR-linearizability) (good term????).
+In case of a failure, once the process is recovered it is "aware" of the failure, in the sense that it runs a designated recovery code.
+In a sense, this recovery code "extends" the interval of the interrupted operation.
+We require that the pending operation linearizes between it's invocation and the end of the recovery code, or aborts.
+In addition, in case of linearization, the returned value of the operation is known to the process.
+Notice that this definition introduce a relaxation which is not intuitive neither trivial - in case that a process operation is complete before the crash, but the response of it is lost due to the cache's lost, we allow the process, upon recovering, extending the interval of that operation using the recovery code (even though the operation was complete), such that at the end of it the process knows the returned value or aborts.
+This gives the process the flexibility to move the linearization point to after the crash, if needed.
+In addition, although the definition allows a process to abort the interrupted operation, it is equivalent to a definition in which a process must complete the operation - once a process knows the operation is aborted, it can reissue it.
+
+RR-linearizability does not specify the progress property required from the implementation.
+One can always use a recoverable mutual exclusion lock at the recovery code in order to sequentially let the processes complete their pending operations.
+Obviously, this solution is not desired when the operation implementation we are recovering is wait-free.
+In this case, we would like the recovery code to be wait-free as well, meaning that a process finish it in a finite number of its own steps in the absence of failure.
+One more point that requires our attention is the possibility of a failure along the recovery code.
+Would it be realistic to assume that such a scenario will not happen?
+Of course, a recovery code that can suffer itself failures is better.
+However, this is not well defined.
+If we allow each failure to run new and different recovery code, we can always use nesting technique.
+A better approach will be such that in the case of failure inside of a recovery code, restarting the recovery code from the beginning is enough.
+In such case, if the recovery code is wait-free in the absence of a failure, we can guarantee that no matter how many failures the process experienced, if given enough time with no failure, it will finish the recovery code (and therefore the pending operation) in a wait-free manner.
+
 \bibliographystyle{abbrv}

### 2017-09-28 article.tex
primitives
recovering read and write

@@
 \usepackage{amsmath}
+\usepackage{algorithm}
+\usepackage[noend]{algpseudocode}
 
@@
 \subsubsection{Recoverable response}
-All the correctness properties discussed above focused on keeping the object or data structure consist in case of a failures.
+All the correctness properties discussed above focused on keeping the object or the data structure consist in case of a failures.
 However, none of which address the problem of the returned value by the failed operation.
 Consider for example a write operation.
-As the operation is atomic, it obviously satisfies strict linearizability - either the write took effect before the crash failure, or that it did not.
-However, once the process is recovered it has no knowledge which of the two is the case, since the acknowledgement returned by the write is saved to the local memory of the process (part of the cache), which is lost due to the failure.
-Assuming the process is in a middle of executing a bigger operation which it desired to complete once it is recovered, it must know whether the write took effect (in which case it can continue and execute the rest of the code), or that it did not (in which case it needs to rewrite).
-Skipping the write is not an option, as it might cause an undesired behaviour which in turn might violate the requirements of the object.
+Since write is atomic it obviously satisfies strict linearizability - either the write took effect before the crash failure, or that it did not.
+However, once the process is recovered it has no knowledge which of the two is the case, since the acknowledgement returned by the write is saved to the local memory of the process (a part of the cache), which is lost due to the failure.
+Assuming the process is in a middle of executing an operation which it desired to complete once it is recovered, it must know whether the write took effect (in which case it can continue and execute the code), or that it did not (in which case it needs to rewrite).
+Skipping the write is not an option, as it might cause an undesired behaviour, which in turn might violate the requirements of the object.
 
-For this case, we present a different approach called Recoverable response linearizability (RR-linearizability) (good term????).
+For this case, we present a different approach called recoverable response linearizability (RR-linearizability) (good term????).
 In case of a failure, once the process is recovered it is "aware" of the failure, in the sense that it runs a designated recovery code.
@@
 However, this is not well defined.
-If we allow each failure to run new and different recovery code, we can always use nesting technique.
-A better approach will be such that in the case of failure inside of a recovery code, restarting the recovery code from the beginning is enough.
-In such case, if the recovery code is wait-free in the absence of a failure, we can guarantee that no matter how many failures the process experienced, if given enough time with no failure, it will finish the recovery code (and therefore the pending operation) in a wait-free manner.
+If we allow each failure to run new and different recovery code, we can use a nesting technique in a way which increase the running time with each failure.
+A better approach will be such that in the case of failure inside of a recovery code, restarting the recovery code, from the beginning any other point, is enough.
+In such case, if the recovery code is wait-free in the absence of a failure, we can guarantee that no matter how many failures the process experienced, once given enough time with no failure, it will finish the recovery code (and therefore the pending operation) in a wait-free manner.
+
+\subsubsection*{Recoverable primitives}
+
+The correctness properties variants discussed earlier do not need to consider primitives, as any atomic operation satisfies any of the properties.
+Unlike it, RR-linearizability is not satisfies by primitives since the respond of the primitive is lost in case of a failure.
+Therefore, there is a need to supply each primitive with an implementation as well as a recovery code.
+In the following section we do so for well known primitives, as well as proving impossibility result for others.
+We assume a failure model of a full-system crash, and address the wait-free progress property.
+The fact that RR-linearizabilty allows us to swift the linearization point of an operation to after the crash is used for the recover after a primitive failure.
+
+\paragraph*{Read}
+
+The process will simply reread the variable upon its recovering. Since read does not affect the rest of the processes, this can be seen as aborting the read preceding the crash, and reissue it.
+
+\paragraph*{Write}
+Write instruction is "wrap" with code such that in case of a failure the extra data will be used for recovering.
+For an instruction writing value $x$ to variable $R$ by process $p$, we provide the following implementation.
+We use the convention of capital letters names for shared memory variables, and small letters for local ones.
+In the following code, $R_p$ is a variable in the memory designated for process $p$.
+
+\begin{algorithm}
+	\caption{Write}\label{euclid}
+	\begin{algorithmic}[1]
+		\Procedure{write operation}{}
+		\State $res \gets R$
+		\State $R_{p} \gets res$
+		\State $R \gets x$
+		\EndProcedure
+
+		\Procedure{recovery code}{}
+		\If {$R_{p} == R$} \Return abort
+		\EndIf
+		\EndProcedure
+	\end{algorithmic}
+\end{algorithm}
+
+For simplicity, we write the recovery code as a single instruction, although it needs to be written as several instructions, as it access two different locations at the memory.
+Since $R_p$ is accessed by $p$ only, the point where $p$ reads $R$ determine the outcome of the recovery code.
+Moreover, in case of a failure at the recovery code, upon recovering the process will execute the recovery code again.
+
+In case of a failure before line 4, the process will abort upon recovering (and reissue the write operation).
+In case of a failure at line 4, the process will execute the recovery code.
+Assuming the if statement at the recovery code does hold, then no process, including $p$, wrote to $R$ after the read of $p$ at line 2, and thus the write operation is aborted and needs to be reissued.
+Otherwise, there must be a process to write to $R$ after the read at line 2.
+In case $p$ write at line 4 took affect before the crash  we linearize it at the point of the write.
+Otherwise there must be a process $q \neq p$ such that $p$ wrote to $R$ in between the two reads of $p$, and we linearize the write of $p$ just before $q$'s write.
+Notice that the write of $q$ covers the write of $p$, and therefore other processes can not distinguish the new execution with the linearized write from the run with no write by $p$.
+
+The above analysis ignores the ABA problem - it might be that $p$ reads the value from $R$, even though there was a write to $R$ in between the two different reads.
+To overcome this problem, we can augment any value written with the writing process's id, and a sequential number (each process will have its own sequential number).
+This way, reading the same value guarantee that no write to $R$ took place between the two different reads.
 

### 2017-10-01 article.tex
TAS
added TAS primitive

@@
 
-The correctness properties variants discussed earlier do not need to consider primitives, as any atomic operation satisfies any of the properties.
+The correctness properties variants discussed earlier does not need to consider primitives, as any atomic operation satisfies any of the properties.
 Unlike it, RR-linearizability is not satisfies by primitives since the respond of the primitive is lost in case of a failure.
 Therefore, there is a need to supply each primitive with an implementation as well as a recovery code.
-In the following section we do so for well known primitives, as well as proving impossibility result for others.
+In the following section we do so for well known primitives, as well as presenting impossibility result for others.
 We assume a failure model of a full-system crash, and address the wait-free progress property.
@@
 
-The process will simply reread the variable upon its recovering. Since read does not affect the rest of the processes, this can be seen as aborting the read preceding the crash, and reissue it.
+The process simply abort the read upon its recovering. Since read does not affect the rest of the processes, this can be done safely.
 
@@
 
-For simplicity, we write the recovery code as a single instruction, although it needs to be written as several instructions, as it access two different locations at the memory.
+For simplicity, we write the recovery code as a single instruction, although it needs to be written as several instructions, as it access two different locations in the memory.
 Since $R_p$ is accessed by $p$ only, the point where $p$ reads $R$ determine the outcome of the recovery code.
-Moreover, in case of a failure at the recovery code, upon recovering the process will execute the recovery code again.
+Moreover, since the recovery code contains only reads, and we already know how to recover reads, the recovery code itself is recoverable.
 
-In case of a failure before line 4, the process will abort upon recovering (and reissue the write operation).
-In case of a failure at line 4, the process will execute the recovery code.
-Assuming the if statement at the recovery code does hold, then no process, including $p$, wrote to $R$ after the read of $p$ at line 2, and thus the write operation is aborted and needs to be reissued.
-Otherwise, there must be a process to write to $R$ after the read at line 2.
-In case $p$ write at line 4 took affect before the crash  we linearize it at the point of the write.
-Otherwise there must be a process $q \neq p$ such that $p$ wrote to $R$ in between the two reads of $p$, and we linearize the write of $p$ just before $q$'s write.
-Notice that the write of $q$ covers the write of $p$, and therefore other processes can not distinguish the new execution with the linearized write from the run with no write by $p$.
+The intuition for the correctness of the code is that if there was a write to $R$ between the two reads of $p$ (at line 2, and at the recovery code), then either this write is by $p$, and we can linearize it at the point where it took affect, or that there was a write by some other process, and we can linearize the write of $p$ just before it, such that we overwrite it, and the rest of the processes can not distinguish between the two situations.
+Therefore, in case of a failure before line 4, $p$ will simply abort upon recovering, and in case of a failure at line 4, $p$ will execute the recovery code.
 
@@
 
+\paragraph*{Test-and-Set}
+We now show that there is no wait-free RR-linearizability implementation for TAS for two processes, assuming the system supply only read/write and TAS primitives.
+We give a high level description for the proof.
+
+Assume by contradiction that there is such an implementation.
+Denote the set of processes by $P=\{p,q\}$.
+First, notice that the code implementing the TAS operation must contain a TAS object $O_1$ which accessed by the two processes and determine their output, otherwise we can solve consensus for two processes using reads and writes only, contradicting to FLP result [*].
+We let both processes run to the point where they both about to access $O_1$, and then perform a TAS one after the other, followed by a crash of $p$ only.
+
+Once $p$ is recover, it needs to execute a recovery code, since the value returned from $O_1$ is lost.
+We look at a solo run of $p$ starting from this point, and consider the possible return values.
+If $p$ returns 1, then we look at the execution where $q$ was the first to access $O_1$ before the crash.
+If $p$ returns 0 or abort, we look at the execution where $q$ is the second to access $O_1$.
+Notice that after the crash, $p$ is not able to distinguish between the different executions, thus it performs the same run after any of these.
+TAS can be seen as a consensus for two processes, the one that returns 1 wins.
+In any case we get a bivalet configuration - both processes want to return the same value (we consider abort as 0), while only one can return 1, while the other must return 0.
+Thus, it must be that both processes access another TAS object $O_2$.
+We can continue and use the same argument, constructing an execution where more and more different TAS objects are accessed.
+
 \bibliographystyle{abbrv}

### 2017-10-02 article.tex
Model 1
model section start

@@
 Correctness property specifies the requirements for an implementation to be considered correct.
-Herlihy and Wing’s linearizability property [*] is widely adopted for shared objects in conventional shared memory models.
+Herlihy and Wing’s \emph{linearizability} property [*] is widely adopted for shared objects in conventional shared memory models.
 Roughly speaking, linearizability requires that operations on objects appear to take effect instantaneously in some serial order, while preserving the happens before relation, i.e., if $op_1$ ends before $op_2$ begins, then $op_1$ should precede $op_2$ in the serial order.
@@
 
-Aguilera and Frølund have proposed strict linearizability as a safety condition for persistent concurrent objects, which treats the crash of a process as a response, either successful or unsuccessful, to the interrupted operation [*].
+Aguilera and Frølund have proposed \emph{strict linearizability} as a safety condition for persistent concurrent objects, which treats the crash of a process as a response, either successful or unsuccessful, to the interrupted operation [*].
 A successful response means that the operation takes effect at some point between its invocation and the crash failure, and an unsuccessful response means that the operation does not take effect at all.
@@
 
-Guerraoui and Levy have proposed persistent atomicity [*].
+Guerraoui and Levy have proposed \emph{persistent atomicity} [*].
 It is similar to strict linearizability, but allows an operation interrupted by a failure to take effect before the subsequent invocation of the same process, possibly after the failure.
-They also proposed transient atomicity, which relaxes this criterion even further and allows an interrupted operation to take effect before the subsequent write response of the same process.
+They also proposed \emph{transient atomicity}, which relaxes this criterion even further and allows an interrupted operation to take effect before the subsequent write response of the same process.
 Both conditions ensure that the state of an object will be consistent in the wake of a crash, but it does not provide locality: correct histories of separate objects, when merged, will not necessarily yield a correct composite history.
@@
 Berryhill et al.
-have proposed an alternative, recoverable linearizability [*], which achieves locality but may sacrifice program order after a crash.
+have proposed an alternative, \emph{recoverable linearizability} [*], which achieves locality but may sacrifice program order after a crash.
 It is a relaxed version of persistent atomicity, which requires the operation to be linearizes or aborts before any subsequent linearization by the pending thread on that same object.
@@
 Under this model, persistent atomicity and recoverable linearizability are indistinguishable (and thus local).
-The term durable linearizability was used to refer to this merged safety condition under the restricted failure model.
+The term \emph{durable linearizability} was used to refer to this merged safety condition under the restricted failure model.
 
@@
 Consider for example a write operation.
-Since write is atomic it obviously satisfies strict linearizability - either the write took effect before the crash failure, or that it did not.
+Since write is atomic it obviously satisfies strict linearizability - either the write took effect before the crash failure, or it did not.
 However, once the process is recovered it has no knowledge which of the two is the case, since the acknowledgement returned by the write is saved to the local memory of the process (a part of the cache), which is lost due to the failure.
@@
 
-For this case, we present a different approach called recoverable response linearizability (RR-linearizability) (good term????).
+For this case, we present a different approach called \emph{recoverable response linearizability} (RR-linearizability) (good term????).
 In case of a failure, once the process is recovered it is "aware" of the failure, in the sense that it runs a designated recovery code.
@@
 This gives the process the flexibility to move the linearization point to after the crash, if needed.
-In addition, although the definition allows a process to abort the interrupted operation, it is equivalent to a definition in which a process must complete the operation - once a process knows the operation is aborted, it can reissue it.
+In addition, although the definition allows a process to abort the interrupted operation, it is equivalent to a definition in which a process must complete the operation - once a process knows the operation was aborted, it can reissue it.
+
+\subsection{Model}
 
 RR-linearizability does not specify the progress property required from the implementation.
-One can always use a recoverable mutual exclusion lock at the recovery code in order to sequentially let the processes complete their pending operations.
+One can use a recoverable mutual exclusion lock at the recovery code in order to sequentially let the processes complete their pending operations.
 Obviously, this solution is not desired when the operation implementation we are recovering is wait-free.

### 2017-10-03 article.tex
Model 2

@@
 \maketitle
-
-\begin{abstract}
-	a
-\end{abstract}
-
-\section{Introduction}
-
-Traditional computer architectures containing both a volatile main memory alongside a non-volatile secondary storage, have led to a separation of program state into operational data stored using in-memory data structures, and recovery data stored using sequential on-disk structures such as transaction logs.
-A system-wide failure, such as a power outage, cause a full lost of the in-memory data structures, which must be reconstructed entirely from the recovery data, making the software system temporarily unavailable.
-As a result, the design of in-memory data structures emphasise disposable constructs optimized for parallelism, in contrast to the structures that hold recovery data, which cannot be discarded upon a failure and which benefit less from parallelism since their performance is limited by the secondary storage.
-
-Current industry projections indicate that non-volatile main memory (NVRAM) will become commonplace over the next few years.
-While the availability of NVRAM suggests the possibility of keeping persistent data in main memory (not just in the file system), the fact that recent updates to registers and cache may be lost during a power failure means that the data in main memory, if not carefully managed, may not be consistent at recovery time.
-Traditional log-based recovery techniques, although can be applied correctly in such systems, fail to take full advantage of the parallelism enabled by allowing processing cores to access recovery data directly using memory operations rather than slow block transfers from secondary storage.
-As a result, harnessing the performance benefits of NVRAM-based platforms requires a careful rethinking of recovery mechanisms.
-
-The study of shared data structures and concurrent programming dates back over a half century ago.
-1965 Dijkstra's pioneering work presented the concept of concurrent programming [*], followed by a series of seminal papers on inter-process communication, wait-free synchronization, and linearizability [*].
-Recent research has paid close attention to asynchronous models, in which there is no bound on the amount of time a process takes to transition to its next step, or to complete a memory operation.
-This assumption captures the behaviour of most modern memory hierarchies, where different media (e.g., L1 cache vs.
-L2 cache vs.
-main memory) incur vastly different and often unpredictable access latencies, as well as the effect of the operating system (e.g., via preemption and interrupts) on the liveness of processes.
-In addition, asynchrony is also related to reliability in a precise way: algorithms that provide non-blocking progress properties (e.g., lock-freedom and wait-freedom) in an asynchronous environment with reliable processes continue to provide the same progress properties if crash failures are introduced.
-Informally speaking, this property holds because a process that crashes permanently at an arbitrary point in the execution of its algorithm is indistinguishable to the other processes from one that is merely very slow.
 
@@
 One more point that requires our attention is the possibility of a failure along the recovery code.
-Would it be realistic to assume that such a scenario will not happen?
+Will it be realistic to assume that such a scenario will not happen?
 Of course, a recovery code that can suffer itself failures is better.
@@
 If we allow each failure to run new and different recovery code, we can use a nesting technique in a way which increase the running time with each failure.
-A better approach will be such that in the case of failure inside of a recovery code, restarting the recovery code, from the beginning any other point, is enough.
+This may also requires infinite memory and code.
+A better approach will be such that in the case of failure inside of a recovery code, restarting the recovery code from some point is enough.
 In such case, if the recovery code is wait-free in the absence of a failure, we can guarantee that no matter how many failures the process experienced, once given enough time with no failure, it will finish the recovery code (and therefore the pending operation) in a wait-free manner.
+
+The standard abstract model is consider in this work.
+The system consist of set of asynchronous unreliable processes $P=\{p_1,\ldots,p_n\}$ that communicate by applying operations on shared \emph{base objects} that support atomic operations such as reads, writes, and read-modify-write primitives.
+
+We assume that a once a process crash, the system knows at which line of the code the fail happened.
+However, there is no indication whether the instruction at that line tool affect before the crash.
+Upon recovering, the controller moves the process to a recovery code according to the line where the crash occur.
+By running this code, the process is aware that a crash happened, as well as at what line of the code was it.
 

### 2017-10-08 article.tex
Model 3
Model done.
to do: references + formal definitions

@@
 The standard abstract model is consider in this work.
-The system consist of set of asynchronous unreliable processes $P=\{p_1,\ldots,p_n\}$ that communicate by applying operations on shared \emph{base objects} that support atomic operations such as reads, writes, and read-modify-write primitives.
+The system consist of a set of asynchronous unreliable processes $P=\{p_1,\ldots,p_n\}$ that communicate by applying operations on shared \emph{base objects} that support atomic operations such as reads, writes, and read-modify-write primitives.
+All base objects are non-volatile, i.e, in case of a crash the values of the base objects is preserved.
+In order to implement a more complex object, such as stack and counter, each process is supplied with access procedures that simulate each operation on the implemented object using operations on base objects.
 
-We assume that a once a process crash, the system knows at which line of the code the fail happened.
-However, there is no indication whether the instruction at that line tool affect before the crash.
-Upon recovering, the controller moves the process to a recovery code according to the line where the crash occur.
-By running this code, the process is aware that a crash happened, as well as at what line of the code was it.
+The failure model considered in this paper is individual.
+Each process may crash, recover, and then continue executing, at any point of the execution.
+Once a process crash, we assume that the system control knows at which line of the process code the failure occurs, or more precisely, at the course of which atomic operation the process failed.
+However, there is no induction whether the atomic operation took affect or not before the crash.
+Each process is also equipped with a recovery code for any possible failure, that is, for any point of the execution it specifies the recovery code in case of a crash at that point.
+Upon recovering, the system control moves the process to the beginning of the appropriate recovery code.
+In this sense, by running the recovery code, the process is aware of the crash.
+At the end of the recovery code, the process goes back to the main code, to the location specified by the recovery code.
+The above model can be implemented by storing the program counter at the non-volatile memory.
+In this case, after each operation on a base object, there should be an operation that updates the value of the program counter.
+This can be done by the system control, and the programmer need not be aware of it.
+Despite doubling the number of accesses to the main memory, this approach reduce the complexity of recovering in case of a crash.
+
+Real world system presents more complications to the abstract model described above.
+In most modern system, each process has its own local volatile cache, which is used to store variables accessed by the process.
+Moreover, in order to improve performance in such systems, writes do not go directly to the shared memory, but rather stored in a buffer at the process's cache, while the system commit the writes later on according to the protocol and model it satisfies.
+Real system thus introduce reordering of instructions, as well as the possibility of losing large amount of data in case of a failure, which makes the recovery procedure harder and more complicate.
+However, the abstract model can be simulated in many of those architectures using special instructions as memory fences and avoid using the cache.
+The cost of such simulation nonetheless, might be very expensive in terms of running time.
+The results presented in this paper implies that even in the simple abstract model the recovery procedure is not trivial. For future work, more complicated and realistic models needs to be studied.
 
@@
 Therefore, there is a need to supply each primitive with an implementation as well as a recovery code.
-In the following section we do so for well known primitives, as well as presenting impossibility result for others.
-We assume a failure model of a full-system crash, and address the wait-free progress property.
-The fact that RR-linearizabilty allows us to swift the linearization point of an operation to after the crash is used for the recover after a primitive failure.
+A system which supply recoverable primitives can be used to implement any object in a recoverable way - in case of a crash the process will simply recover the last atomic operation it executed before the crash.
+In case the operation is aborted, the process will reissue it.
+After that, it can continue and execute the rest of its code safely.
+This observation focus our attention to implementation of primitives in a recoverable manner.
+
+In the following section we present RR-liniarize implementations for well known primitives, as well as presenting impossibility result for others.
+We focus our attention on bounded wait-free implementations, that is, the number of steps a process takes when executing the recovery code in the absence of a failure is finite and bounded by a known constant (may be a function of n, the number of processes in the system), regardless of the other processes steps and the failures the process experienced before.
+In addition, we would like the recovery code to use a finite number of variables.
+The fact that RR-linearizabilty allows us to swift the linearization point of an operation to after the crash is used to recover after a primitive failure.
 
@@
 \paragraph*{Test-and-Set}
-We now show that there is no wait-free RR-linearizability implementation for TAS for two processes, assuming the system supply only read/write and TAS primitives.
+
+Test-and-Set (TAS) is a known object supported as a primitive in many modern architectures.
+A TAS object contains a binary value which initialized to 0, and support a single atomic operation TAS which set the value of the object to 1 and returns the old value.
+
+We now show that there is no bounded wait-free RR-linearizability implementation for TAS for two processes, assuming the system supply only read/write and TAS primitives.
 We give a high level description for the proof.
+The construction also implies that any wait-free implementation need to use an unbounded number of TAS objects.
 

### 2017-10-08 article.tex
references
references added

@@
 \usepackage[noend]{algpseudocode}
+\usepackage{cite}
 
@@
 
-\subsection{Correctness properties}
+\subsection{Correctness properties} \label{Correctness properties}
 Correctness property specifies the requirements for an implementation to be considered correct.
-Herlihy and Wing’s \emph{linearizability} property [*] is widely adopted for shared objects in conventional shared memory models.
+Herlihy and Wing’s \emph{linearizability} property \cite{DBLP:journals/toplas/HerlihyW90} is widely adopted for shared objects in conventional shared memory models.
 Roughly speaking, linearizability requires that operations on objects appear to take effect instantaneously in some serial order, while preserving the happens before relation, i.e., if $op_1$ ends before $op_2$ begins, then $op_1$ should precede $op_2$ in the serial order.
@@
 
-Aguilera and Frølund have proposed \emph{strict linearizability} as a safety condition for persistent concurrent objects, which treats the crash of a process as a response, either successful or unsuccessful, to the interrupted operation [*].
+Aguilera and Fr{\o}lund have proposed \emph{strict linearizability} as a safety condition for persistent concurrent objects, which treats the crash of a process as a response, either successful or unsuccessful, to the interrupted operation \cite{Aguilera2003StrictLA}.
 A successful response means that the operation takes effect at some point between its invocation and the crash failure, and an unsuccessful response means that the operation does not take effect at all.
@@
 
-Guerraoui and Levy have proposed \emph{persistent atomicity} [*].
+Guerraoui and Levy have proposed \emph{persistent atomicity} \cite{DBLP:conf/icdcs/GuerraouiL04}.
 It is similar to strict linearizability, but allows an operation interrupted by a failure to take effect before the subsequent invocation of the same process, possibly after the failure.
@@
 Berryhill et al.
-have proposed an alternative, \emph{recoverable linearizability} [*], which achieves locality but may sacrifice program order after a crash.
+have proposed an alternative, \emph{recoverable linearizability} \cite{DBLP:conf/opodis/BerryhillGT15}, which achieves locality but may sacrifice program order after a crash.
 It is a relaxed version of persistent atomicity, which requires the operation to be linearizes or aborts before any subsequent linearization by the pending thread on that same object.
@@
 Izraelevitz et al.
-considered a real-world failure model, in which processes are assumed to fail together, as part of a full-system crash.
+considered a real-world failure model, in which processes are assumed to fail together, as part of a full-system crash \cite{DBLP:conf/wdag/IzraelevitzMS16}.
 Under this model, persistent atomicity and recoverable linearizability are indistinguishable (and thus local).
@@
 
-For this case, we present a different approach called \emph{recoverable response linearizability} (RR-linearizability) (good term????).
+For this case, we present a different approach called \emph{recoverable response linearizability} (RR-linearizability).
 In case of a failure, once the process is recovered it is "aware" of the failure, in the sense that it runs a designated recovery code.
@@
 In addition, in case of linearization, the returned value of the operation is known to the process.
-Notice that this definition introduce a relaxation which is not intuitive neither trivial - in case that a process operation is complete before the crash, but the response of it is lost due to the cache's lost, we allow the process, upon recovering, extending the interval of that operation using the recovery code (even though the operation was complete), such that at the end of it the process knows the returned value or aborts.
+Notice that this definition introduce a relaxation which is not intuitive neither trivial - in case that a process operation took affect before the crash, but the response of it is lost due to the cache's lost, we allow the process, upon recovering, extending the interval of that operation using the recovery code, such that at the end of it the process knows the returned value or aborts.
 This gives the process the flexibility to move the linearization point to after the crash, if needed.
 In addition, although the definition allows a process to abort the interrupted operation, it is equivalent to a definition in which a process must complete the operation - once a process knows the operation was aborted, it can reissue it.
+
+\iffalse
+Progress properties are formally defined using histories.
+We follow [*] for these definitions.
+There are four types of steps: (1) an invocation step, denoted $(INV, p, X, op)$, represents the invocation by process $p$ of operation $op$ on implemented object $X$; (2) a response step, denoted $(RES, p, X, ret)$, represents the completion by process $p$ of the last operation it invoked on object $X$, with response $ret$; (3) a crash step, denoted $(CRASH, p)$, denotes a crash of process $p$; (4) a recovery step, denoted $(REC, p)$, denotes a recovery of process $p$;
+
+A history $H$ is a sequence of steps, possibly involving multiple processes and implemented objects.
+For a given history $H$, we denote by $H \mid p$ the projection of $H$ onto the steps of process $p$.
+Similarly, we denote by $H \mid O$ the projection of $H$ onto the steps involving implemented object $O$.
+We adopt the convention that both $H \mid p$ and $H \mid O$ retain all crash steps in $H$.
+A response step is \emph{matching} with respect to an invocation step $s$ by a process $p$ on object $X$ in a history $H$ if it is the first response step by $p$ on $X$ that follows $s$ in $H$, and it occurs before p’s next invocation (if any) in $H$.
+\fi
 
@@
 RR-linearizability does not specify the progress property required from the implementation.
-One can use a recoverable mutual exclusion lock at the recovery code in order to sequentially let the processes complete their pending operations.
+One can use a recoverable mutual exclusion lock \cite{DBLP:conf/podc/GolabR16,DBLP:conf/podc/GolabH17} at the recovery code in order to sequentially let the processes complete their pending operations.
 Obviously, this solution is not desired when the operation implementation we are recovering is wait-free.
 In this case, we would like the recovery code to be wait-free as well, meaning that a process finish it in a finite number of its own steps in the absence of failure.
-One more point that requires our attention is the possibility of a failure along the recovery code.
+Another point that requires our attention is the possibility of a failure along the recovery code.
 Will it be realistic to assume that such a scenario will not happen?
@@
 The system consist of a set of asynchronous unreliable processes $P=\{p_1,\ldots,p_n\}$ that communicate by applying operations on shared \emph{base objects} that support atomic operations such as reads, writes, and read-modify-write primitives.
-All base objects are non-volatile, i.e, in case of a crash the values of the base objects is preserved.
+All base objects are non-volatile, i.e, in case of a crash the values stored in the base objects are preserved.
 In order to implement a more complex object, such as stack and counter, each process is supplied with access procedures that simulate each operation on the implemented object using operations on base objects.
@@
 Each process may crash, recover, and then continue executing, at any point of the execution.
-Once a process crash, we assume that the system control knows at which line of the process code the failure occurs, or more precisely, at the course of which atomic operation the process failed.
+Once a process crash, we assume that the system control knows at which line of the code the failure occurs, or more precisely, at the course of which atomic operation the process failed.
 However, there is no induction whether the atomic operation took affect or not before the crash.
@@
 The above model can be implemented by storing the program counter at the non-volatile memory.
-In this case, after each operation on a base object, there should be an operation that updates the value of the program counter.
+In this case, after each operation on a base object, there should be an operation updating the value of the program counter.
 This can be done by the system control, and the programmer need not be aware of it.
@@
 Moreover, in order to improve performance in such systems, writes do not go directly to the shared memory, but rather stored in a buffer at the process's cache, while the system commit the writes later on according to the protocol and model it satisfies.
-Real system thus introduce reordering of instructions, as well as the possibility of losing large amount of data in case of a failure, which makes the recovery procedure harder and more complicate.
-However, the abstract model can be simulated in many of those architectures using special instructions as memory fences and avoid using the cache.
-The cost of such simulation nonetheless, might be very expensive in terms of running time.
+Real system thus introduce reordering of instructions, as well as the possibility of losing large amount of data in case of a failure, which makes the recovery procedure harder and more complicated.
+However, the abstract model can be simulated in many of those architectures, by using special instructions as memory fences and by avoid using the cache.
+The cost of such simulation, nonetheless, might be very expensive in terms of running time.
 The results presented in this paper implies that even in the simple abstract model the recovery procedure is not trivial. For future work, more complicated and realistic models needs to be studied.
@@
 
-The correctness properties variants discussed earlier does not need to consider primitives, as any atomic operation satisfies any of the properties.
-Unlike it, RR-linearizability is not satisfies by primitives since the respond of the primitive is lost in case of a failure.
+The correctness properties variants discussed in section \ref{Correctness properties} does not need to consider primitives, as any atomic operation satisfies any of the properties.
+Unlike it, RR-linearizability is not satisfied by primitives since the respond of the primitive is lost in case of a failure.
 Therefore, there is a need to supply each primitive with an implementation as well as a recovery code.
@@
 In case the operation is aborted, the process will reissue it.
-After that, it can continue and execute the rest of its code safely.
+After that, it can continue and execute the remaining code safely.
 This observation focus our attention to implementation of primitives in a recoverable manner.
 
-In the following section we present RR-liniarize implementations for well known primitives, as well as presenting impossibility result for others.
-We focus our attention on bounded wait-free implementations, that is, the number of steps a process takes when executing the recovery code in the absence of a failure is finite and bounded by a known constant (may be a function of n, the number of processes in the system), regardless of the other processes steps and the failures the process experienced before.
+In the following section we present RR-linearizable implementations for well known primitives, as well as presenting impossibility result for others.
+We focus our attention on bounded wait-free implementations, that is, the number of steps a process takes when executing the recovery code in the absence of a failure is finite and bounded by a known constant (may be a function of n, the number of processes in the system), regardless of the other processes steps and the failures the process experienced so far.
 In addition, we would like the recovery code to use a finite number of variables.
@@
 Test-and-Set (TAS) is a known object supported as a primitive in many modern architectures.
-A TAS object contains a binary value which initialized to 0, and support a single atomic operation TAS which set the value of the object to 1 and returns the old value.
+A TAS object contains a binary value initialized to 0, and support a single atomic operation TAS which set the value of the object to 1 and returns the old value.
 
-We now show that there is no bounded wait-free RR-linearizability implementation for TAS for two processes, assuming the system supply only read/write and TAS primitives.
+We now show that there is no bounded wait-free RR-linearizable implementation for TAS even for two processes, assuming the system support only read, write and TAS primitives.
 We give a high level description for the proof.
@@
 If $p$ returns 0 or abort, we look at the execution where $q$ is the second to access $O_1$.
-Notice that after the crash, $p$ is not able to distinguish between the different executions, thus it performs the same run after any of these.
+Notice that after the crash, $p$ does not able to distinguish between the different executions, thus it performs the same run after any of these.
 TAS can be seen as a consensus for two processes, the one that returns 1 wins.
@@
 
-\bibliographystyle{abbrv}
-\bibliography{references}
+\bibliographystyle{plain}
+\bibliography{references} {}
 

### 2017-10-24 article.tex
cas algorithm
research goals

@@
 In addition, in case of linearization, the returned value of the operation is known to the process.
-Notice that this definition introduce a relaxation which is not intuitive neither trivial - in case that a process operation took affect before the crash, but the response of it is lost due to the cache's lost, we allow the process, upon recovering, extending the interval of that operation using the recovery code, such that at the end of it the process knows the returned value or aborts.
+Notice that this definition introduce a relaxation which is not intuitive neither trivial - in case that a process operation took affect before the crash, but the response of it is lost due to the cache's lost, we allow the process, upon recovering, to extend the interval of that operation using the recovery code, such that at the end of it the process knows the returned value or aborts.
 This gives the process the flexibility to move the linearization point to after the crash, if needed.
@@
 
+\paragraph*{Compare-and-Swap}
+
+Compare-and-Swap (CAS)...
+
+\begin{algorithm}
+	\caption{Compare-and-Swap}\label{euclid}
+	\begin{algorithmic}[1]
+		\Procedure{cas operation}{}
+		\State $<id,val> \gets C$
+		\If {$val \neq old$}
+		\State \Return false
+		\EndIf
+		\State $S[i] \gets <id,val>$
+		\State $res \gets C.cas(<id,val>, <i,new>)$
+		\If {$res = true$}
+		\State $R[id][i] \gets val$
+		\EndIf
+		\State $S[i] \gets null$
+		\EndProcedure
+
+		\Procedure{recovery code}{}
+		\State Read C, S[], R[i][]
+		\If {$<id,val>$ appears in C or S[], or $val$ appears in R[i][]}
+		\State \Return true
+		\Else {}
+		\State \Return false
+		\EndIf
+		\EndProcedure
+	\end{algorithmic}
+\end{algorithm}
+
 \paragraph*{Test-and-Set}

### 2017-10-24 research goals.tex
research goals

--- 2017-10-08 article.tex
@@
-\documentclass{article}
+\documentclass[11pt]{article}
+\usepackage{times,cite}
+\usepackage{graphicx}
+\usepackage{amsmath}
+\usepackage{amssymb}
 
-\usepackage{amsmath}
-\usepackage{algorithm}
-\usepackage[noend]{algpseudocode}
-\usepackage{cite}
+\setlength{\evensidemargin}{0.2in}
+\setlength{\oddsidemargin}{0.2in}
+\setlength{\marginparwidth}{0.6in}
+\setlength{\textwidth}{6.20in}
+\newlength{\figwidth}
+\setlength{\figwidth}{6.0in}
+\setlength{\textheight}{8.00in}
+\setlength{\topmargin}{0.25in}
+\setlength{\headheight}{0.15in}
+%\setlength{\headsep}{0in}
+\renewcommand{\baselinestretch}{1.2}
 
-%opening
-\title{ISF Grant}
-\author{}
+\usepackage{wrapfig}
+\usepackage[vlined,ruled,linesnumbered,noresetcount]{algorithm2e}
+\DontPrintSemicolon
+
+\markright{PIs: Hagit Attiya and Danny Hendler, application: 380/18}\pagestyle{myheadings}
+\renewcommand{\thepage}{}
+\pagestyle{empty}
+
+\newcommand{\remove}[1]{}
+
+\makeatletter
+\newcommand{\removelatexerror}{\let\@latex@error\@gobble}
+\makeatother
 
@@
 
-\maketitle
+\section{Detailed Description of the Proposed Research}
 
-\subsection{Correctness properties} \label{Correctness properties}
-Correctness property specifies the requirements for an implementation to be considered correct.
-Herlihy and Wing’s \emph{linearizability} property \cite{DBLP:journals/toplas/HerlihyW90} is widely adopted for shared objects in conventional shared memory models.
-Roughly speaking, linearizability requires that operations on objects appear to take effect instantaneously in some serial order, while preserving the happens before relation, i.e., if $op_1$ ends before $op_2$ begins, then $op_1$ should precede $op_2$ in the serial order.
-Linearizability has no notion of an aborted or failed operation, and requires that a process finish one operation before it invokes the next.
-The criteria according to which implementations that support crash-recovery failures are correct remains unclear.
+We now describe the key goals of our research in more detail. \\
 
-Aguilera and Fr{\o}lund have proposed \emph{strict linearizability} as a safety condition for persistent concurrent objects, which treats the crash of a process as a response, either successful or unsuccessful, to the interrupted operation \cite{Aguilera2003StrictLA}.
-A successful response means that the operation takes effect at some point between its invocation and the crash failure, and an unsuccessful response means that the operation does not take effect at all.
-Strict linearizability preserves both locality and program order.
-However, they proved that it precludes the implementation of some wait-free objects for certain machine models - there is no wait-free implementations of multi-reader single-writer (MRSW) registers from single-reader single-writer (SRSW) registers under strict linearizability.
-
-Guerraoui and Levy have proposed \emph{persistent atomicity} \cite{DBLP:conf/icdcs/GuerraouiL04}.
-It is similar to strict linearizability, but allows an operation interrupted by a failure to take effect before the subsequent invocation of the same process, possibly after the failure.
-They also proposed \emph{transient atomicity}, which relaxes this criterion even further and allows an interrupted operation to take effect before the subsequent write response of the same process.
-Both conditions ensure that the state of an object will be consistent in the wake of a crash, but it does not provide locality: correct histories of separate objects, when merged, will not necessarily yield a correct composite history.
-
-Berryhill et al.
-have proposed an alternative, \emph{recoverable linearizability} \cite{DBLP:conf/opodis/BerryhillGT15}, which achieves locality but may sacrifice program order after a crash.
-It is a relaxed version of persistent atomicity, which requires the operation to be linearizes or aborts before any subsequent linearization by the pending thread on that same object.
-
-Izraelevitz et al.
-considered a real-world failure model, in which processes are assumed to fail together, as part of a full-system crash \cite{DBLP:conf/wdag/IzraelevitzMS16}.
-Under this model, persistent atomicity and recoverable linearizability are indistinguishable (and thus local).
-The term \emph{durable linearizability} was used to refer to this merged safety condition under the restricted failure model.
-
-\subsubsection{Recoverable response}
-All the correctness properties discussed above focused on keeping the object or the data structure consist in case of a failures.
+\noindent \emph{{\bf Research goal 1:} Identify the right specification
+	of NVMM.}\\
+All the correctness properties discussed in previous section focus on keeping the object or the data structure consist in case of a failures.
 However, none of which address the problem of the returned value by the failed operation.
@@
 Since write is atomic it obviously satisfies strict linearizability - either the write took effect before the crash failure, or it did not.
-However, once the process is recovered it has no knowledge which of the two is the case, since the acknowledgement returned by the write is saved to the local memory of the process (a part of the cache), which is lost due to the failure.
+However, once the process is recovered it has no knowledge which of the two is the case, since the acknowledgement returned by the write is saved to the local memory of the process, which is lost due to the failure.
 Assuming the process is in a middle of executing an operation which it desired to complete once it is recovered, it must know whether the write took effect (in which case it can continue and execute the code), or that it did not (in which case it needs to rewrite).
@@
 In addition, in case of linearization, the returned value of the operation is known to the process.
-Notice that this definition introduce a relaxation which is not intuitive neither trivial - in case that a process operation took affect before the crash, but the response of it is lost due to the cache's lost, we allow the process, upon recovering, extending the interval of that operation using the recovery code, such that at the end of it the process knows the returned value or aborts.
+Notice that this definition introduce a relaxation which is not intuitive neither trivial - in case that a process operation took affect before the crash, but the response of it is lost due to the crash, we allow the process, upon recovering, to extend the interval of that operation using the recovery code, such that at the end of it the process knows the returned value or aborts.
 This gives the process the flexibility to move the linearization point to after the crash, if needed.
@@
 
-\iffalse
-Progress properties are formally defined using histories.
-We follow [*] for these definitions.
-There are four types of steps: (1) an invocation step, denoted $(INV, p, X, op)$, represents the invocation by process $p$ of operation $op$ on implemented object $X$; (2) a response step, denoted $(RES, p, X, ret)$, represents the completion by process $p$ of the last operation it invoked on object $X$, with response $ret$; (3) a crash step, denoted $(CRASH, p)$, denotes a crash of process $p$; (4) a recovery step, denoted $(REC, p)$, denotes a recovery of process $p$;
-
-A history $H$ is a sequence of steps, possibly involving multiple processes and implemented objects.
-For a given history $H$, we denote by $H \mid p$ the projection of $H$ onto the steps of process $p$.
-Similarly, we denote by $H \mid O$ the projection of $H$ onto the steps involving implemented object $O$.
-We adopt the convention that both $H \mid p$ and $H \mid O$ retain all crash steps in $H$.
-A response step is \emph{matching} with respect to an invocation step $s$ by a process $p$ on object $X$ in a history $H$ if it is the first response step by $p$ on $X$ that follows $s$ in $H$, and it occurs before p’s next invocation (if any) in $H$.
-\fi
-
-\subsection{Model}
-
-RR-linearizability does not specify the progress property required from the implementation.
-One can use a recoverable mutual exclusion lock \cite{DBLP:conf/podc/GolabR16,DBLP:conf/podc/GolabH17} at the recovery code in order to sequentially let the processes complete their pending operations.
-Obviously, this solution is not desired when the operation implementation we are recovering is wait-free.
-In this case, we would like the recovery code to be wait-free as well, meaning that a process finish it in a finite number of its own steps in the absence of failure.
-Another point that requires our attention is the possibility of a failure along the recovery code.
-Will it be realistic to assume that such a scenario will not happen?
-Of course, a recovery code that can suffer itself failures is better.
-However, this is not well defined.
-If we allow each failure to run new and different recovery code, we can use a nesting technique in a way which increase the running time with each failure.
-This may also requires infinite memory and code.
-A better approach will be such that in the case of failure inside of a recovery code, restarting the recovery code from some point is enough.
-In such case, if the recovery code is wait-free in the absence of a failure, we can guarantee that no matter how many failures the process experienced, once given enough time with no failure, it will finish the recovery code (and therefore the pending operation) in a wait-free manner.
-
-The standard abstract model is consider in this work.
+There are also various flavours for the model.
+We first consider the standard abstract model, a simple well-studied model.
 The system consist of a set of asynchronous unreliable processes $P=\{p_1,\ldots,p_n\}$ that communicate by applying operations on shared \emph{base objects} that support atomic operations such as reads, writes, and read-modify-write primitives.
@@
 In order to implement a more complex object, such as stack and counter, each process is supplied with access procedures that simulate each operation on the implemented object using operations on base objects.
+Each process is equipped with a local volatile memory.
+We assume no caches, thus writes goes directly to the main memory.
+In addition, upon a process crash, we assume that the system control knows at which line of the code the failure occurs, or more precisely, at the course of which atomic operation the process failed.
+However, there is no induction whether the atomic operation took affect or not before the crash.
 
-The failure model considered in this paper is individual.
-Each process may crash, recover, and then continue executing, at any point of the execution.
-Once a process crash, we assume that the system control knows at which line of the code the failure occurs, or more precisely, at the course of which atomic operation the process failed.
-However, there is no induction whether the atomic operation took affect or not before the crash.
-Each process is also equipped with a recovery code for any possible failure, that is, for any point of the execution it specifies the recovery code in case of a crash at that point.
-Upon recovering, the system control moves the process to the beginning of the appropriate recovery code.
-In this sense, by running the recovery code, the process is aware of the crash.
-At the end of the recovery code, the process goes back to the main code, to the location specified by the recovery code.
-The above model can be implemented by storing the program counter at the non-volatile memory.
-In this case, after each operation on a base object, there should be an operation updating the value of the program counter.
-This can be done by the system control, and the programmer need not be aware of it.
-Despite doubling the number of accesses to the main memory, this approach reduce the complexity of recovering in case of a crash.
+Another point to consider is how to model failures.
+Clearly, the whole system can fail, but we will consider a more relaxed model in which individual processes can also fail.
 
-Real world system presents more complications to the abstract model described above.
-In most modern system, each process has its own local volatile cache, which is used to store variables accessed by the process.
-Moreover, in order to improve performance in such systems, writes do not go directly to the shared memory, but rather stored in a buffer at the process's cache, while the system commit the writes later on according to the protocol and model it satisfies.
-Real system thus introduce reordering of instructions, as well as the possibility of losing large amount of data in case of a failure, which makes the recovery procedure harder and more complicated.
-However, the abstract model can be simulated in many of those architectures, by using special instructions as memory fences and by avoid using the cache.
-The cost of such simulation, nonetheless, might be very expensive in terms of running time.
-The results presented in this paper implies that even in the simple abstract model the recovery procedure is not trivial. For future work, more complicated and realistic models needs to be studied.
+What is the right progress condition?
 
-\subsubsection*{Recoverable primitives}
-
-The correctness properties variants discussed in section \ref{Correctness properties} does not need to consider primitives, as any atomic operation satisfies any of the properties.
-Unlike it, RR-linearizability is not satisfied by primitives since the respond of the primitive is lost in case of a failure.
-Therefore, there is a need to supply each primitive with an implementation as well as a recovery code.
-A system which supply recoverable primitives can be used to implement any object in a recoverable way - in case of a crash the process will simply recover the last atomic operation it executed before the crash.
+\noindent \emph{{\bf Research goal 2:} Derive algorithms, lower bounds and complexity tradeoffs}\\
+First, we look for RR-linearized implementation for base object.
+This direction attract our focus thanks to the following observation: a system which supply recoverable primitives can be used to implement any object in a recoverable way.
+In case of a crash the process will simply recover the last atomic operation it executed before the crash.
 In case the operation is aborted, the process will reissue it.
 After that, it can continue and execute the remaining code safely.
-This observation focus our attention to implementation of primitives in a recoverable manner.
 
-In the following section we present RR-linearizable implementations for well known primitives, as well as presenting impossibility result for others.
-We focus our attention on bounded wait-free implementations, that is, the number of steps a process takes when executing the recovery code in the absence of a failure is finite and bounded by a known constant (may be a function of n, the number of processes in the system), regardless of the other processes steps and the failures the process experienced so far.
-In addition, we would like the recovery code to use a finite number of variables.
-The fact that RR-linearizabilty allows us to swift the linearization point of an operation to after the crash is used to recover after a primitive failure.
-
-\paragraph*{Read}
-
-The process simply abort the read upon its recovering. Since read does not affect the rest of the processes, this can be done safely.
-
-\paragraph*{Write}
-Write instruction is "wrap" with code such that in case of a failure the extra data will be used for recovering.
-For an instruction writing value $x$ to variable $R$ by process $p$, we provide the following implementation.
-We use the convention of capital letters names for shared memory variables, and small letters for local ones.
-In the following code, $R_p$ is a variable in the memory designated for process $p$.
-
-\begin{algorithm}
-	\caption{Write}\label{euclid}
-	\begin{algorithmic}[1]
-		\Procedure{write operation}{}
-		\State $res \gets R$
-		\State $R_{p} \gets res$
-		\State $R \gets x$
-		\EndProcedure
-
-		\Procedure{recovery code}{}
-		\If {$R_{p} == R$} \Return abort
-		\EndIf
-		\EndProcedure
-	\end{algorithmic}
-\end{algorithm}
-
-For simplicity, we write the recovery code as a single instruction, although it needs to be written as several instructions, as it access two different locations in the memory.
-Since $R_p$ is accessed by $p$ only, the point where $p$ reads $R$ determine the outcome of the recovery code.
-Moreover, since the recovery code contains only reads, and we already know how to recover reads, the recovery code itself is recoverable.
-
-The intuition for the correctness of the code is that if there was a write to $R$ between the two reads of $p$ (at line 2, and at the recovery code), then either this write is by $p$, and we can linearize it at the point where it took affect, or that there was a write by some other process, and we can linearize the write of $p$ just before it, such that we overwrite it, and the rest of the processes can not distinguish between the two situations.
-Therefore, in case of a failure before line 4, $p$ will simply abort upon recovering, and in case of a failure at line 4, $p$ will execute the recovery code.
-
-The above analysis ignores the ABA problem - it might be that $p$ reads the value from $R$, even though there was a write to $R$ in between the two different reads.
-To overcome this problem, we can augment any value written with the writing process's id, and a sequential number (each process will have its own sequential number).
-This way, reading the same value guarantee that no write to $R$ took place between the two different reads.
-
-\paragraph*{Test-and-Set}
+Some base objects have an RR-linearzied implementation.
+For example, read has a simple implementation - the process will simply abort the read upon its recovering.
+Since read does not affect the rest of the processes, this can be done safely.
+However, not every primitive has an RR-linearized implementation.
 
@@
 A TAS object contains a binary value initialized to 0, and support a single atomic operation TAS which set the value of the object to 1 and returns the old value.
-
 We now show that there is no bounded wait-free RR-linearizable implementation for TAS even for two processes, assuming the system support only read, write and TAS primitives.
@@
 TAS can be seen as a consensus for two processes, the one that returns 1 wins.
-In any case we get a bivalet configuration - both processes want to return the same value (we consider abort as 0), while only one can return 1, while the other must return 0.
+In any case we get a bivalet configuration - both processes want to return the same value (we consider abort as 0), while only one can return 1 and the other must return 0.
 Thus, it must be that both processes access another TAS object $O_2$.
@@
 
-\bibliographystyle{plain}
-\bibliography{references} {}
+Future research will focus on implementing additional primitives, as write and compare-and-swap (which have an RR-linearized implementation).
+In addition, the complexity of those implementations is another research direction.
+Is there a tradeoff between normal operation and recovery?
+Can one achieve the best on both?
+Is there a tradeoff between space and time for such implementations?
+
+Another research direction is looking for implementation for special object.
+Although the approach of using RR-linearized base objects is universal, it might be that certain objects as stack or counter can be implemented directly more efficiently without the use of RR-linearized base object, or with reducing the use of such base objects.
+
+For RME, there are still several open problems. Is $log n / loglog n$ RMRs optimal for CC model? What can be done in the DSM model?
+
+\noindent \emph{{\bf Research goal 3:} Incorporate novel architecture aspects.}\\
+The model presented in goal 1 is very simple and abstract.
+The model is used in order to investigate the right specification for NVRAM systems.
+As described earlier, even under this simple model there are impossibility results, and non trivial implementation.
+However, the model neglect some of the complexities of real world systems.
+Current architectures contains volatile cache in addition to main memory.
+In such systems, write does not goes directly to the main memory, but rather to a cached copy.
+Therefore, in case of a failure, the write can be lost due to the cache lost.
+Moreover, such systems introduce reordering of instruction.
+Fence instruction is used to prevent such reordering.
+While it guarantees that all writes preceding it will become visible, in practice it does not necessary means writes goes to the main memory.
+In such case, a crash after a fence instruction still may cause a lost of writes.
+
+A more realistic model should take the above architectures and the complications they present into account.
+Also, real world system may contain a combinations of NVRAM and ordinary memory, and a new model needs to be devised for such systems.
+One more practical property that NVRAM has is a big difference between the cost of a read and the cost of write.
+So far the research has neglected this property, treating both operations as the same cost.
+Weather we can reduce the number of writes in certain implementation, even in the cost of additional reads, is another research direction.
 
